\section{Conclusions}\label{chapter-04:section:conclusion}

In this chapter, we have presented the implementation of the RxInfer framework, which is a
reactive message passing-based inference framework.
RxInfer introduces reactive continual inference and
user-friendly languages for model and constraints specification.
These languages are designed to resemble mathematical equations and make the specification of
complex inference procedures simpler and less error-prone.
The reactivity of the implementation brings several advantages to the inference process, as
discussed in Chapter~\ref{chapter-03}.
The framework is capable of performing inference in sophisticated probabilistic models and
supports infinite data streams as part of its core design.
RxInfer aims to lower the bar for researchers to experiment with complex probabilistic models
and sophisticated inference procedures, providing a user-friendly \ac{api} that hides the
associated complexity of model and constraints specification, as well as the execution of the
inference.
Additionally, RxInfer is highly extensible and modular, allowing users to define custom nodes,
message update rules, and approximation methods at its core, providing flexibility and
customization options.

RxInfer is an open source framework that is available to everyone and is actively maintained
by BIASlab at the Eindhoven University of Technology.
In the next chapter, we will present an experimental evaluation of the RxInfer framework in
large-scale signal processing and related applications.
We will demonstrate that RxInfer can comfortably scale to models with hundreds of thousands of
random variables and perform real-time and continual inference in complex and hierarchical
models.
