% !TEX root = ../main.tex

\chapter{Acknowledgments}

% Bert, Wouter Kouw
To my supervisor and friend \textbf{Bert de Vries}. 
Meeting you was a pivotal moment in my life, and you stand out as one of the most interesting people I have ever known. I still vividly remember our first casual encounter and our very first walk to the office when you kindly suggested a good coffeeshop. Even though I have never tried their coffee, I appreciated the recommendation. Thank you for giving me the opportunity to pursue my Ph.D. in BIASlab and work alongside exceptional people. Being your student was a privilege, as you have a unique talent for bringing together outstanding teams. 
Your guidance, belief in my potential, and the freedom you gave me have been invaluable throughout this journey.

% Other members of the committee
To the members of the defense committee \textbf{Carel Peeters}, \textbf{Eric-Jan Wagenmakers}, \textbf{Erik Meijer}, and \textbf{Ruud van Sloun}. I sincerely appreciate the time and effort you invested in reviewing my work and providing valuable feedback. Thanks to \textbf{Huib Visser} for being the chairman during my defense. A special thanks to my co-promotor \textbf{Wouter Kouw}. I am truly impressed with your ability to engage in insightful discussions and offer constructive feedback, even when you had no opportunity to sleep for more than 2 hours.

% BIASlab group
% - Bart
% - Wouter Nuijten
% - Thijs
% - Hoang
% - Chengfeng
% - Magnus
% - Mykola
% - Raaja
% - Omer
% - Sepideh
% - Tim
% - Marco
% - Joris
% - Mattia

To \textbf{BIASlab}, a group of passionate and wonderful people.
\textbf{Bart van Erp}, I have never met a person so well organized as you. 
I think you could easily finish two Ph.D.s in two years while still being a part-time accountant and part-time researcher in LazyDynamics. 
\textbf{Thijs van de Laar}, thank you for creating ForneyLab. I can only truly appreciate how difficult it was to put such
a complex mathematical theory in a package using Julia version 0.4, where you could probably finish your cup of coffee waiting for a single line plot.
\textbf{Magnus Koudahl}, thank you for always finding (intentionally and unintentionally) new ways to break RxInfer. 
Constant feedback from you only helped to improve the ecosystem. You are also a great musician and I really
enjoyed our occasional jamming sessions.
\textbf{Semih Akbayrak}, it was a great pleasure working with you and learning from your experience.
It felt like you had an answer to every question possible.
\textbf{Wouter Nuijten}, I am glad you joined our lab.  
You are the person who not only comprehended my coding ideas, even when they were a bit scattered, 
but also managed to implement them before I could fully articulate them.
Many thanks to other members of the group, \textbf{Hoang}, \textbf{Chengfeng}, \textbf{Joris}, \textbf{Mykola}, \textbf{Raaja}, \textbf{Sepideh}, \textbf{Tim}, \textbf{Marco}, \textbf{Omer} and \textbf{Mattia}, for their feedback and a great working environment. 

% Jan Bergmans
% Signal Process Group
% Secretariat (Emerald)
Thanks to \textbf{Jan Bergmans} for creating a wonderful work atmosphere. I would also like to express my gratitude to the secretariat of the Signal Processing Systems group at TU/e and, in particular, to \textbf{Emerald Clump-Busser}, for their help with various nontechnical matters.

% GN sponsors
Many thanks to \textbf{GN Advanced Science} for sponsoring my work.

% Study before my PhD 
% INM, Igor Konshin & Yuri Vasilevskii
To my former supervisor \textbf{Igor Konshin}. I am deeply grateful for your kindness and invaluable mentorship during the early stages of my scientific journey. In such a short period of time, I learned numerous valuable lessons from you, which continue to help me in my career. I extend my gratitude to \textbf{Yuri Vasilevskii} for the opportunity to engage in research in various corners of the globe.

% IBCH, Mikhail Shugay, Dima Bolotin and Mitya Chudakov
To my mentor \textbf{Mikhail Shugay}. The experience I gained from working with you is beyond words. 
I will always cherish your guidance and the insightful projects we have undertaken together.
Additionally, I extend my thanks to \textbf{Dima Bolotin} and \textbf{Mitya Chudakov}, 
with whom I have the pleasure of working in a positive and friendly environment.

% Teachers 
% Bart Nederhand 
To my drumteacher \textbf{Bart Nederhand}. Thank you for guiding me into the drumming world. 
Your lessons have been a constant source of inspiration.
It all began during my very first encounter with drumsticks at your trial lesson. 
From that moment on, my desire to play the drums has never faded, but has only grown stronger. 

% Blanka van Es
Tijdens mijn PhD-reis heb ik niet alleen wiskunde of drumspelen geleerd, maar ook een beetje Nederlands. 
Hartelijk dank, \textbf{Blanka van Es}, voor het delen van een stukje Nederlandse cultuur. 
Ik heb echt genoten van elk gesprek.

% JuliaCon folks
% - Gareth
% - Mathijs
% - Jorge
To a group of Julia enthusiasts from \textbf{Julia Eindhoven Meetup}, including \textbf{Gareth}, \textbf{Matthijs} and \textbf{Jorge}.
It is truly motivating to belong to a group of individuals who casually (and voluntarily) organize the largest technical events dedicated to the Julia programming language in the Eindhoven region.

% Skydiving folks
% - Jan
% - Eleonora
% - Merlijn
% - Joost
% - Tom
% - Felix
% - Alexandru
To my fellow skydiving buddies at \textbf{Skydive Flanders}, including \textbf{Eleonora}, \textbf{Jan}, \textbf{Eloy}, \textbf{Joost}, \textbf{Merlijn}, \textbf{Tom}, \textbf{Felix}, \textbf{Alex}, and \textbf{Yulia}. We have gone through many ups and downs. Thank you for sharing the same passion for the skies and being supportive members of the club. 
Special thanks to \textbf{Anton}, who casually walked into my office and left me with little choice but to sign up for my first skydiving experience. I wish I had signed up earlier.

% Friends abroad
% Bulat Akhmetshin
% Ilya Boldyrev
% Ilya Belov
% Andrey Chaplygin
% Kamilla
% Anton Mitrofanov
% Valentina Shumovskaya
% Tanya Lebedeva
% Olya Cvetkova
% Pasha Talamanov
% Natalya Teplyakova
% Lukaschuk Anton.
% Vlad Pimanov
% Denis Anuprienko
% Stas Abukhovich
To my friends abroad. 
Thank you \textbf{Bulat}, \textbf{Ilya}, and \textbf{Andrey}. The multitude of memories and adventures we have shared together deserves its own book. Thank you \textbf{Vlad}, \textbf{Kamilla}, \textbf{Anton}, \textbf{Ilya}, \textbf{Pasha}, \textbf{Natasha}, \textbf{Olya}, \textbf{Valentina}, \textbf{Denis}, \textbf{Stas}, \textbf{Yulya}, and \textbf{Tanya} for staying in touch with me. 
I wish I have had as many opportunities to meet you as I would like.

% Anton & Kirill Alexeev
% Sasha
% Semih & Melike
% Alyona
% Branislav
% Yunus
% Alp
% Sara
% Martin
% Lev
% Marco
% Jiali
% Omar & Daniela
% Ozan
% Anton & Diana
% Ivan
During my PhD I have met so many wonderful people in the Netherlands. Thank you \textbf{Anton}, \textbf{Kirill}, \textbf{Sasha}, \textbf{Semih}, \textbf{Melike}, \textbf{Alëna}, \textbf{Branislav}, \textbf{Yunus}, \textbf{Alp}, \textbf{Sara}, \textbf{Martin}, \textbf{Lev}, \textbf{Marco}, \textbf{Jiali}, \textbf{Omar}, \textbf{Daniela}, \textbf{Ozan}, \textbf{Anton}, \textbf{Diana} and \textbf{Ivan}. You made my journey more enjoyable and my home is always a welcoming place for you.

% Eindhoven Friends
% Anastassia
To \textbf{Anastassia Imatra}.
Thank you for all the moments we have shared.
Despite your belief that people do not change, I am confident that you have played a significant role in shaping who I am today. 

% Ismail
To \textbf{Ismail Șenöz}. You are not just an exceptional researcher, but also an awesome friend. 
Big thanks for taking the time to teach me how math really works. 
I cannot imagine getting through this complicated journey without your steady help and patience.
Watching you tackle tough math problems was always impressive, but hearing you play the guitar was even more beautiful. 
I loved every jam session we had (except for that one funny hiccup).
Special thanks to your lovely family, who warmly welcomed me in the best places in Turkey.

% Gleb & Katya
To \textbf{Gleb Nazarikov}. We first met in the Netherlands; surprising, considering how many mutual friends we had. Yet, it was an instant connection, and we embarked on the entire Ph.D. journey together. I can honestly say that I couldn't have asked for a better housemate. In fact, I cannot even imagine someone better, and I doubt such a person exists. Thank you for tolerating my drumming practice sessions. However, the mystery of those wildboars remains unsolved, and perhaps it always will. Special thanks to \textbf{Katya Malysheva} who always enthusiastically joined Gleb and me in playing board games, even when Buienradar suggested not to.

% Albert
To \textbf{Albert Podusenko}. Anyone who has moved to another country knows how difficult it really is.
Albert, you became my first friend in this unfamiliar world, my best friend.
I was happy when you chose to stay in the Netherlands, even though the decision was not entirely yours. 
Your constant belief meant a lot to me and I could never finish this journey without your friendship.
I appreciate that, despite all the difficulties you have had in your life, you never stopped supporting me.
And please, abandon your attempts to permanently injure yourself. 
It is not like a coin; one does not need a lot of observations to infer that bones have the remarkable ability to heal.

To my cousin \textbf{Vanya}, my uncle \textbf{Andrey} and my aunt \textbf{Olya}. 
Although our meetings are not very frequent, your presence and kindness undoubtedly enrich my life. 
Thank you for your influence during the moments we share.

% - Grandparents mom
% - Grandparents dad
\begin{otherlanguage*}{russian}
\textsf{Дорогие дедушка \textsf{\textbf{Саша}} и бабушка \textsf{\textbf{Нина}}.
Ваши добрые слова и поддержка были для меня настоящей опорой и вдохновением.
Невзирая на расстояние, ваш вклад в мою жизнь остается неизгладимым. 
Дорогие дедушка \textsf{\textbf{Костя}} и бабушка \textsf{\textbf{Рая}}.
Ваши любовь и тепло остаются со мной и я всегда буду помнить вас с благодарностью в сердце.
}
\end{otherlanguage*}

% - Sasha + Vanya + Olya/Andrey
To my brother \textbf{Sasha}. Thanks for always giving in to me when we had fights.
I don't want to imagine how my life would have turned out if you hadn't told Dad about AESC.
Since then, I have always followed your steps and you are my closest companion, my second soul.
And today, even though we live on different sides of the planet, your influence and motivation still reach me.

% - Alliance
% - Parents
Last but certainly not least, I want to express my immense gratitude to my lovely parents, \textbf{Elena} and \textbf{Vladislav}.
Words cannot adequately convey what truly wonderful people you are.
Even our feline friend \textbf{Alliance} does undoubtedly agree with me without saying a single word. 
Through your love and care, you miraculously transformed the once grumpy and closed creature
into the most lovely cat in the world, eagerly running to you at every opportunity.
For me, reaching this point would not have been possible without your unwavering support. 
I am grateful for all the effort, sacrifices, and love.
Thank you for always giving me more than I could ever ask for, 
for letting me go, explore the vast unknown world, and become a person I hope you can be proud of.



\begin{flushright}
-- Dmitry
\end{flushright}

